---
title: "Sports Crime 2026 Report"
author: "Addina Rahaman and Alexie DaCosta"
date: '`r Sys.time()`'
output:
  html_document:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: 3
---

# Replication

## Load Packages
```{r setup, include=FALSE}
library(here)
library(scales)
library(tidyverse)
library(cfbfastR)
library(broom)
library(stringr)
library(lubridate)
library(ggeffects)
```

## Load Data
```{r}
offenses <- read_csv(here('offense_segment.csv')) 
batches <- read_csv(here('nibrs_batch_header_1991_2024.csv')) %>% 
  filter(grepl('200[0-5]', year))
```

## Organize and join

### Offenses Data
Filter and organize offenses within the programs of interest and only Group A offenses (assaults, vandalism)
```{r}
university_ori_lookup <- tribble(
  ~university, ~city_name, ~ori,
  'the university of akron', 'akron', 'OH0772800',
  'iowa state university', 'ames', 'IA0850100',
  'the university of michigan', 'ann arbor', 'MI8190300',
  'ohio university', 'athens', 'OH0050100',
  'the university of texas at austin', 'austin', 'TX2270500',
  'virginia polytechnic institute and state university', 'blacksburg', 'VA0600100',
  'boise state university', 'boise', 'ID0010100',
  'clemson university', 'clemson', 'SC0390600', 
  'united states air force academy', 'colorado springs', 'CO0210100',
  'the university of south carolina', 'columbia', 'SC0400600',
  'the ohio state university', 'columbus', 'OH0252700',
  'the university of north texas', 'denton', 'TX0611100',
  'michigan state university', 'east lansing', 'MI3336400',
  'university of arkansas', 'fayetteville', 'AR0720400',
  'marshall university', 'huntington', 'WV0060200',
  'the university of iowa', 'iowa city', 'IA0520400',
  'arkansas state university', 'jonesboro', 'AR0160500',
  'western michigan university', 'kalamazoo', 'MI3949900',
  'the university of kansas', 'lawrence', 'KS0230100',
  'utah state university', 'logan', 'UT0031600',
  'texas tech university', 'lubbock', 'TX1520600',
  'west virginia university', 'morgantown', 'WV0310100',
  'the university of idaho', 'moscow', 'ID0290500', 
  'eastern michigan university', 'mount pleasant', 'MI3759900',
  'middle tennessee state university', 'murfreesboro', 'TN0750100',
  'brigham young university', 'provo', 'UT0250600'
)

batches_filtered <- university_ori_lookup %>% 
  left_join(batches, by=c('ori', 'city_name'))

```

```{r}
offenses_filtered <- batches_filtered %>% 
  left_join(offenses) %>% 
  drop_na(year)

# All the assault and vandalism cases
offenses_filtered <- offenses_filtered %>%
 filter(grepl('assault|vandalism', ucr_offense_code, ignore.case=TRUE))
```

### Game Schedule Data
Get the game schedules for 2000-2005
```{r}
game_schedule_raw <- c()
years <- 2000:2005

for (year in years) {
  i = year %% 2000 + 1
  game_schedule_raw[[i]] = espn_cfb_schedule(
    year = year
  )
}

names(game_schedule_raw) = c('2000', '2001', '2002', '2003', '2004', '2005')

game_schedule <- bind_rows(game_schedule_raw, .id='year')

game_schedule <- game_schedule %>% 
  mutate(
    home_team_name = case_when(
      home_team_id == "2711" ~ "Broncos2", # Boise and western michigan have the same team name so im replacing western michigan with Broncos2
      TRUE ~ home_team_name
    )
  )
```

### Join and Aggregate
Join offenses data and schedule data and aggregate to get the number of offenses per agency day, and respective game/team information
```{r}
university_team_lookup <- tribble(
  ~university, ~team_name,
  'the university of akron', 'Zips',
  'iowa state university', 'Cyclones',
  'the university of michigan', 'Wolverines',
  'ohio university', 'Bobcats', 
  'the university of texas at austin', 'Longhorns',
  'virginia polytechnic institute and state university', 'Hokies',
  'boise state university', 'Broncos',
  'clemson university', 'Tigers',
  'united states air force academy', 'Falcons',
  'the university of south carolina', 'Gamecocks',
  'the ohio state university', 'Buckeyes',
  'the university of north texas', 'Mean Green',
  'michigan state university', 'Spartans',
  'university of arkansas', 'Razorbacks',
  'marshall university', 'Thundering Herd',
  'the university of iowa', 'Hawkeyes',
  'arkansas state university', 'Red Wolves', 
  'western michigan university', 'Broncos2',
  'the university of kansas', 'Jayhawks',
  'utah state university', 'Aggies', 
  'texas tech university', 'Red Raiders',
  'west virginia university', 'Mountaineers',
  'the university of idaho', 'Vandals',
  'eastern michigan university', 'Eagles',
  'middle tennessee state university', 'Blue Raiders',
  'brigham young university', 'Cougars' 
)

# create a column
offenses_filtered <- offenses_filtered %>% 
  mutate(
    offense_type = case_when(
      (grepl('assault', ucr_offense_code)) ~ 'assault',
      (grepl('vandalism', ucr_offense_code)) ~ 'vandalism'
    )
  )

# create a column that classifies by no game, home game, and away game
agency_days <- offenses_filtered %>% 
  filter((grepl('^200[0-5]-', incident_date))) %>%
  group_by(university, ori, incident_date, offense_type) %>%
  summarize(
    number_of_offenses=n()
  )

game_schedule_filtered <- game_schedule %>% 
  select(year, matchup, season, game_date, home_team_name, away_team_name)

agency_days_and_programs <- agency_days %>% 
  left_join(university_team_lookup)

agency_offenses_and_games <- agency_days_and_programs %>% 
  left_join(game_schedule_filtered, by=c('incident_date' = 'game_date')) %>% 
  mutate(
    game_type = case_when(
      (team_name == home_team_name) ~ 'home_game',
      (team_name == away_team_name) ~ 'away_game',
      TRUE ~ 'No game' # since we joined only by incident date and game date, it could be possible that neither home_team_name or away_team_name corresponds with team_name. in that case, the team was not involved at all
    )
  ) %>% 
  group_by(ori, incident_date, offense_type, number_of_offenses) %>% 
  summarize(
    actual_game_type = if (any(game_type == 'home_game')) { 'home_game' } 
                      else if (any(game_type == 'away_game')) { 'away_game' }
                      else { 'no_game' }
  ) %>% 
  rename(game_type=actual_game_type)
```

## Plotting 

### Appendix Table 1
Statistics on assaults per agency day on (1) all days (2) when there are no games (3) when there are home games, and (4) when there are away games
```{r}
# appendix 1 
appendix_one_game <- agency_offenses_and_games %>% 
  group_by(game_type, offense_type) %>% 
  summarize(
    quantile_0.25 = quantile(number_of_offenses, 0.25, na.rm=TRUE),
    quantile_0.50 = quantile(number_of_offenses, 0.50, na.rm=TRUE),
    quantile_0.75 = quantile(number_of_offenses, 0.75, na.rm=TRUE),
    quantile_0.90 = quantile(number_of_offenses, 0.90, na.rm=TRUE),
    sd = sd(number_of_offenses), 
    avg = mean(number_of_offenses),
    .groups = 'drop'
  ) 

appendix_one_all_day <- agency_offenses_and_games %>% 
  group_by(offense_type) %>% 
  summarize(
    quantile_0.25 = quantile(number_of_offenses, 0.25, na.rm=TRUE),
    quantile_0.50 = quantile(number_of_offenses, 0.50, na.rm=TRUE),
    quantile_0.75 = quantile(number_of_offenses, 0.75, na.rm=TRUE),
    quantile_0.90 = quantile(number_of_offenses, 0.90, na.rm=TRUE),
    sd = sd(number_of_offenses), 
    avg = mean(number_of_offenses),
    .groups = 'drop'
  ) %>% 
  mutate(game_type = 'all_days')

appendix_one <- bind_rows(appendix_one_game, appendix_one_all_day)

appendix_one %>% 
  ggplot(aes(x=game_type, y=avg, ymin=avg-sd, ymax=avg+sd)) +
  geom_pointrange() + 
  geom_errorbar() + 
  facet_wrap(~ offense_type)
```

### Table 2 
Plot the distribution of games and observed offenses by day of week
```{r}
# table 2 (distribution of games by day of week)

agency_offenses_and_games_by_day <- agency_offenses_and_games %>%
  mutate(day_of_week = weekdays(incident_date))
 
table2_obs <- agency_offenses_and_games_by_day  %>%
  group_by(day_of_week) %>%
  summarize(num_observations = sum(number_of_offenses))

table2_games <- game_schedule_filtered %>% 
  filter(home_team_name %in% university_team_lookup$team_name | away_team_name %in% university_team_lookup$team_name) %>% 
  mutate(day_of_week = weekdays(game_date)) %>% 
  group_by(day_of_week) %>% 
  summarize(num_games = n())
 
table2 <- inner_join(table2_games, table2_obs, by='day_of_week')

total <- table2 %>% 
  summarize(
    total_games = sum(num_games), 
    total_observations = sum(num_observations)
  ) %>% 
  mutate(day_of_week='Total') %>% 
  rename(num_games=total_games, num_observations=total_observations)

table2 <- bind_rows(table2, total)

table2 %>% 
  pivot_longer(
    cols=starts_with('num_'),
    names_to='type',
    values_to='stat'
  ) %>%
  ggplot(aes(x=day_of_week, y=stat)) +
  geom_bar(stat = 'identity') + 
  facet_wrap(~ type, scales='free')

```

### Descriptive Figure - Saturdays
Mean and standard error of number of offenses on Saturday (overall).
```{r}
# saturday plot
agency_offenses_and_games %>% 
  mutate(day_of_week=weekdays(incident_date)) %>%
  filter(day_of_week=='Saturday') %>%
  group_by(game_type, offense_type) %>% 
  summarize(
    count = n(),
    avg=mean(number_of_offenses),
    se=sd(number_of_offenses) / sqrt(count)
  ) %>% 
  ggplot(aes(x = game_type, y=avg, ymin=avg-se, ymax=avg+se)) +
  geom_pointrange() +
  geom_errorbar() + 
  labs(
    x = 'Game Type',
    y = 'Number of Offense Reports',
    title = 'Saturday Offense by Agency Days',
  ) +
  facet_wrap(~ offense_type)
```

### Descriptive Figure - Per team
Mean and standard error of offenses on Saturday (each university).
```{r}
# per team plots

# both offense types in one graph (cramped)
agency_offenses_and_games %>% 
  mutate(day_of_week=weekdays(incident_date)) %>%
  filter(day_of_week=='Saturday') %>%
  left_join(university_ori_lookup) %>% 
  group_by(university, game_type, offense_type) %>% 
  summarize(
    count = n(),
    avg=mean(number_of_offenses),
    se=sd(number_of_offenses) / sqrt(count)
  ) %>% 
  ggplot(aes(x = game_type, y=avg, ymin=avg-se, ymax=avg+se, color=offense_type)) +
  geom_pointrange() +
  geom_errorbar() +
  facet_wrap(~ university) + 
  labs(
    x = 'Game Type',
    y = 'Number of Offense Reports',
    title = 'Saturday Offense by Agency Days',
    color = 'Offense Type'
  )

# only assault reports per uni
agency_offenses_and_games %>% 
  mutate(day_of_week=weekdays(incident_date)) %>%
  filter(day_of_week=='Saturday', offense_type=='assault') %>%
  left_join(university_ori_lookup) %>% 
  group_by(university, game_type) %>% 
  summarize(
    count = n(),
    avg=mean(number_of_offenses),
    se=sd(number_of_offenses) / sqrt(count)
  ) %>% 
  ggplot(aes(x = game_type, y=avg, ymin=avg-se, ymax=avg+se)) +
  geom_pointrange() +
  geom_errorbar() +
  facet_wrap(~ university) + 
  labs(
    x = 'Game Type',
    y = 'Number of Assault Reports',
    title = 'Saturday Assaults by Agency Days'
  )

# only vandalism reports per uni
agency_offenses_and_games %>% 
  mutate(day_of_week=weekdays(incident_date)) %>%
  filter(day_of_week=='Saturday', offense_type=='vandalism') %>%
  left_join(university_ori_lookup) %>% 
  group_by(university, game_type) %>% 
  summarize(
    count = n(),
    avg=mean(number_of_offenses),
    se=sd(number_of_offenses) / sqrt(count)
  ) %>% 
  ggplot(aes(x = game_type, y=avg, ymin=avg-se, ymax=avg+se)) +
  geom_pointrange() +
  geom_errorbar() +
  facet_wrap(~ university) + 
  labs(
    x = 'Game Type',
    y = 'Number of Vandalism Reports',
    title = 'Saturday Vandalisms by Agency Days'
  )


```

## Regression - Table 3

### Re-organizing data
```{r}
# psuedo code: lm(daily_offense_count ~ home_game + away_game + agency_FE)
library(MASS)
 
# Hot Encode Game Type
factors <- model.matrix(~ game_type - 1, data = agency_offenses_and_games)
colnames(factors) <- sub("^game_type", "", colnames(factors))
 
# Horizontally stack data back to df
model_data <- bind_cols(agency_offenses_and_games, factors)
 
# Add features
model_data <- model_data %>%
mutate(
  day_of_week = wday(incident_date),
  month = month(incident_date),
  year = year(incident_date)
)


```

### Model Outputs
We did both linear regression models and negative binomial models. We ran 
each type of model on (1) both offense types combined (2) only assault offense types
(3) only vandalism offense types. For each, we also computed the % increases compared
to no game respective to the offense type and model type. 

```{r}

# Overall Model - Linear Regression
mod <- lm(number_of_offenses ~ home_game + away_game + as.factor(ori) + as.factor(day_of_week) + as.factor(month) + as.factor(year),
data = model_data)
 
summary(mod)

# Calculate the percent change
(coef(mod)['home_game'] / mean(model_data$number_of_offenses)) * 100
#home_game 
  #10.4358 
 
# Overall Model - Negative Binomial Regression
mod2 <- glm.nb(number_of_offenses ~ home_game + away_game + as.factor(ori) + as.factor(day_of_week) + as.factor(month) + as.factor(year),
data = model_data)
 
summary(mod2)
 
# Calculate the percent change
( exp(coef(mod2)['home_game']) - 1 ) * 100
# home_game 
# 10.43503 
 
# Linear Regression by Assault
mod_a <- lm(number_of_offenses ~ home_game + away_game + as.factor(ori) + as.factor(day_of_week) + as.factor(month) + as.factor(year),
data = model_data_assault)
 
summary(mod_a)

 # Calculate the percent change
(coef(mod_a)['home_game'] / mean(model_data_assault$number_of_offenses)) * 100
# home_game 
# -0.7868275 
 
# Negative Binomial Regression by Assault
mod_a2 <- glm.nb(number_of_offenses ~ home_game + away_game + as.factor(ori) + as.factor(day_of_week) + as.factor(month) + as.factor(year),
data = model_data_assault)
 
summary(mod_a2)
 
# Calculate the percent change
( exp(coef(mod_a2)['home_game']) - 1) * 100
#home_game 
# 1.487655 
 
# Linear Regression by Vandalism
mod_v <- lm(number_of_offenses ~ home_game + away_game + as.factor(ori) + as.factor(day_of_week) + as.factor(month) + as.factor(year),
data = model_data_vandalism)
 
summary(mod_v)

# Calculate the percent change
 (coef(mod_v)['home_game'] / mean(model_data_vandalism$number_of_offenses)) * 100
# home_game 
# 20.60766 
 
# Negative Binomial Regression by Vandalism
mod_v2 <- glm.nb(number_of_offenses ~ home_game + away_game + as.factor(ori) + as.factor(day_of_week) + as.factor(month) + as.factor(year),
data = model_data_vandalism)
 
summary(mod_v2)
 
# Calculate the percent change
( exp(coef(mod_v2)['home_game']) - 1 ) * 100
# home_game 
# 16.93446
```

### Graphing Regression Outputs
```{r}
coef_df <- bind_rows(
  tidy(mod2, conf.int = TRUE) %>%
    filter(term == "home_game") %>%
    mutate(model = "Overall"),

  tidy(mod_a2, conf.int = TRUE) %>%
    filter(term == "home_game") %>%
    mutate(model = "Assault"),

  tidy(mod_v2, conf.int = TRUE) %>%
    filter(term == "home_game") %>%
    mutate(model = "Vandalism")
) %>%
  mutate(
    pct_effect = (exp(estimate) - 1) * 100,
    pct_low = (exp(conf.low) - 1) * 100,
    pct_high = (exp(conf.high) - 1) * 100
  )

ggplot(coef_df,
       aes(x = model, y = pct_effect)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = pct_low,
                    ymax = pct_high),
                width = .15) +
  geom_hline(yintercept = 0,
             linetype = "dashed") +
  labs(
    x = "",
    y = "Percent Change in Offenses",
    title = "Estimated Home Game Effect"
  ) +
  theme_minimal()
```

# Extension 

## Data Organization

### Legal Betting Lookup Table
```{r}

# url used: https://www.cbssports.com/betting/news/u-s-sports-betting-where-all-50-states-stand-on-legalizing-online-sports-betting-sites-proposed-legislation/

legal_betting_lookup <- tribble(
  ~state_id, ~launch_date,
  'AL', NA,
  'AK', NA, 
  'AZ', as.Date('2021-09-09'),
  'AR', as.Date('2022-03-04'),
  'CA', NA,
  'CO', as.Date('2020-05-01'),
  'CT', as.Date('2021-10-19'),
  'DE', as.Date('2023-12-26'),
  'FL', as.Date('2023-11-07'),
  'GA', NA, 
  'HI', NA, 
  'ID', NA,
  'IL', as.Date('2020-06-18'),
  'IN', as.Date('2019-09-01'),
  'IA', as.Date('2019-08-15'),
  'KS', as.Date('2022-09-01'),
  'KY', as.Date('2023-09-28'),
  'LA', as.Date('2022-01-28'),
  'ME', as.Date('2023-11-03'),
  'MD', as.Date('2022-11-23'),
  'MA', as.Date('2023-03-10'),
  'MI', as.Date('2021-01-22'),
  'MN', NA,
  'MS', NA,
  'MO', NA, # legalized after 2024
  'MT', as.Date('2020-03-01'),
  'NE', NA,
  'NV', as.Date('2010-01-01'),
  'NH', NA,
  'NJ', NA,
  'NM', NA,
  'NY', NA,
  'NC', as.Date('2024-3-11'),
  'ND', NA,
  'OH', as.Date('2023-01-01'),
  'OK', NA,
  'OR', NA,
  'PA', as.Date('2019-05-01'),
  'RI', NA, 
  'SC', NA,
  'SD', NA,
  'TN', NA, 
  'TX', NA,
  'UT', NA, 
  'VT', as.Date('2024-01-11'),
  'VA', NA,
  'WA', as.Date('2020-03-25'),
  'WV', as.Date('2018-08-30'),
  'WI', NA, 
  'WY', as.Date('2021-09-01')

  'WY', as.Date('2021-09-01')
)


```

### Replicate Earlier Table for 2015-2024
```{r}
# load data
offense_segment_2015_2024 <- read_csv('offense_segment_2015_2024.csv')
batches_2015_2024 <- read_csv(here('nibrs_batch_header_1991_2024.csv')) %>% 
  filter(grepl('201[5-9]|202[0-4]', year))

extension_years <- 2015:2024
game_schedule_2015_2024_raw <- c()

for (year in extension_years) {
  i = year %% 2015 + 1
  game_schedule_2015_2024_raw[[i]] = espn_cfb_schedule(
    year = year
  )
}

names(game_schedule_2015_2024_raw) = c('2015', '2016', '2017', '2018', '2019', '2020', '2021', '2022', '2023', '2024')

game_schedule_2015_2024 <- bind_rows(game_schedule_2015_2024_raw, .id='year')

game_schedule_2015_2024 <- game_schedule_2015_2024 %>% 
  mutate(
    home_team_name = case_when(
      home_team_id == "2711" ~ "Broncos2", # Boise and western michigan have the same team name so im replacing western michigan with Broncos2
      TRUE ~ home_team_name
    )
  )

gs_2023 <- espn_cfb_schedule(year=2023)
gs_2023 <- gs_2023 %>% 
  mutate(
    year='2023'
  )
gs_2024 <- espn_cfb_schedule(year=2024)
gs_2024 <- gs_2024 %>% 
  mutate(
    year='2024'
  )

# brute forcing 2023 and 2024 because they were having problems
game_schedule_2015_2024 <- bind_rows(game_schedule_2015_2024, gs_2023, gs_2024)
```

```{r}
# clean data for only programs of interest
university_ori_state_lookup <- university_ori_lookup %>% 
  mutate(state_id=str_sub(ori, 1, 2))

batches_2015_2024_filtered <- university_ori_state_lookup %>% 
  left_join(batches_2015_2024, by=c('ori', 'city_name'))

offenses_2015_2024_filtered <- batches_2015_2024_filtered %>% 
  left_join(offense_segment_2015_2024)

# filter out assault and vandalism 
offenses_2015_2024_filtered <- offenses_2015_2024_filtered %>% 
  filter(grepl('assault|vandalism', ucr_offense_code, ignore.case=TRUE)) %>% 
  mutate(
    offense_type = case_when(
      (grepl('assault', ucr_offense_code)) ~ 'assault',
      (grepl('vandalism', ucr_offense_code)) ~ 'vandalism'
    )
  )

```

```{r}
# aggregate data 

agency_days_2015_2024 <- offenses_2015_2024_filtered %>% 
  filter((grepl('^201[5-9]-|202[0-4]-', incident_date))) %>%
  group_by(university, ori, incident_date, offense_type) %>%
  summarize(
    number_of_offenses=n()
  ) 

game_schedule_2015_2024_filtered <- game_schedule_2015_2024 %>% 
  select(year, matchup, season, game_date, home_team_name, away_team_name)

agency_days_and_programs_2015_2024 <- agency_days_2015_2024 %>% 
  left_join(university_team_lookup)

agency_offenses_and_games_2015_2024 <- agency_days_and_programs_2015_2024 %>% 
  left_join(game_schedule_2015_2024_filtered, by=c('incident_date' = 'game_date')) %>% 
  mutate(
    game_type = case_when(
      (team_name == home_team_name) ~ 'home_game',
      (team_name == away_team_name) ~ 'away_game',
      TRUE ~ 'No game' # since we joined only by incident date and game date, it could be possible that neither home_team_name or away_team_name corresponds with team_name. in that case, the team was not involved at all
    )
  ) %>% 
  group_by(ori, incident_date, offense_type, number_of_offenses) %>% 
  summarize(
    actual_game_type = if (any(game_type == 'home_game')) { 'home_game' } 
                      else if (any(game_type == 'away_game')) { 'away_game' }
                      else { 'no_game' }
  ) %>% 
  rename(game_type=actual_game_type) %>% 
  filter(game_type != 'away_game') %>% 
   mutate(
    state_id=str_sub(ori, 1, 2)
  )

```

```{r}
# add a legal_betting column and treatement column
extension_df <- agency_offenses_and_games_2015_2024

extension_df <- extension_df %>% 
  left_join(legal_betting_lookup, by='state_id') %>% 
  mutate(
    legal_betting = case_when(
      (is.na(launch_date)) ~ 0,
      (incident_date >= launch_date) ~ 1, 
      TRUE ~ 0
    ),
    treatment = ifelse(is.na(launch_date), 0, 1)
  )

# quick sanity check
test <- temp_df %>% 
  filter(state_id == 'AR') %>% 
  filter(incident_date <= launch_date) %>% 
  head(n=5)

```

### Feature Engineering for Extension Data 
```{r}
 
# Hot Encode Game Type
factors <- model.matrix(~ game_type - 1, data = extension_df)
colnames(factors) <- sub("^game_type", "", colnames(factors))
 
# Horizontally stack data back to df
model_data <- bind_cols(extension_df, factors)
 
# Add features
model_data <- model_data %>%
mutate(
  day_of_week = wday(incident_date),
  month = month(incident_date),
  year = year(incident_date)
)

factors(extension_df$game_type)
days2024 <- seq(as.Date('2024-01-01'), as.Date('2024-12-31'))
extension_2024 <- extension_df %>% 
  filter(incident_date %in% days2024)
  
extension_2024 %>% 
  mutate(
    treatment_type = ifelse(treatment == 1, 'Legalized States', 'Never Legalized States')
  ) %>% 
  group_by(offense_type, treatment_type, game_type) %>% 
  summarize(
    count = n(),
    avg=mean(number_of_offenses),
    se=sd(number_of_offenses)/sqrt(count)
  ) %>% 
  ggplot(aes(x=game_type, y=avg, ymin=avg-se, ymax=avg+se)) +
  geom_pointrange() +
  geom_errorbar() +
  facet_wrap(offense_type ~ treatment_type) + 
  labs(
    x = 'Game Type',
    y = 'Number of Offense Reports',
    title = 'Mean Offense Reports for No Game Days and Home Game Days Based on Legalized Betting in 2024'
  )

```

### Plot 2 - Time Series 2015-2024
```{r}
treatment_df <- extension_df %>% 
  filter(treatment == 1)

treatment_states <- unique(treatment_df$state_id) 

num_states <- 1:7 
names(num_states) <- treatment_states
launch_years <- c()

for (state in treatment_states) {
  i <- num_states[state]
  launch_year <- treatment_df %>% 
    filter(state_id == state) %>% 
    mutate(legal_betting_year = year(launch_date)) %>% 
    pull(legal_betting_year)

  launch_years[i] <- launch_year[1]
}

names(launch_years) <- treatment_states
launch_years_df <- enframe(launch_years, name = "state_id", value = "xintercept")


treatment_df %>% 
  mutate(year = year(incident_date)) %>%
  group_by(year, offense_type, game_type) %>% 
  summarize(
    avg_offenses=mean(number_of_offenses),
    .groups = "drop"
  ) %>% 
  pivot_wider(
    names_from=game_type,
    values_from=avg_offenses
  ) %>% 
  mutate(
    premium=home_game-no_game
  ) %>% 
  ggplot(aes(x=year, y=premium, color=offense_type)) + 
  geom_line() + 
  geom_vline(
    data = launch_years_df, 
    aes(xintercept = xintercept), 
    linetype = 'dashed'
  ) + 
  labs(
    title="Premium in Assaults and Vandalism from 2015-2024 by States of Interest",
    x='Years',
    y='Premium'
  ) +
  facet_wrap(~ state_id)

```

### Extension Data: Feature Engineering for Regression Model
```{r}
# Hot Encode Game Type
factors <- model.matrix(~ game_type - 1, data = extension_df)
colnames(factors) <- sub("^game_type", "", colnames(factors))
 
# Horizontally stack data back to df
ext_model_data <- bind_cols(extension_df, factors)
 
# Add features
ext_model_data <- ext_model_data %>%
mutate(
  day_of_week = wday(incident_date),
  month = month(incident_date),
  year = year(incident_date)
)

# Hot Encode ORI and Calendar Features
ext_model_data$ori <- factor(ext_model_data$ori)
ext_model_data$day_of_week <- factor(ext_model_data$day_of_week)
ext_model_data$month <- factor(ext_model_data$month)
ext_model_data$year <- factor(ext_model_data$year)

```

### Predictive Modeling for Extension Data 
```{r}
extension_model <- glm.nb(number_of_offenses ~  + home_game + legal_betting + home_game:legal_betting + 
ori + day_of_week + month + year,
data = ext_model_data)

summary(extension_model)

beta_home <- coef(extension_model)["home_game"]
beta_int  <- coef(extension_model)["home_game:legal_betting"]

# What is the effect of a home game after legalization
pct_change <- (exp(beta_home + beta_int) - 1) * 100
pct_change

# After legalization, home games are associated with 3.7% more offenses than non-home-game days.
```

```{r}
pred <- ggpredict(
  extension_model,
  terms = c("legal_betting", "home_game")
)

pred$group <- factor(
  pred$group,ss  git switch -c rescue-detached
  levels = c("0", "1"),
  labels = c("No", "Yes")
)

plot(pred) +
  labs(
    title = "Predicted Offenses by Legalized Betting and Home Game Status",
    x = "Legalized Betting",
    y = "Predicted Number of Offenses",
    color = "Home Game"
  ) 
```
